### 2022-12-28 quarto/papers.qmd
update

---
title: "Papers"
---

## 2022

* Z. Zhang, **A. Zhang**, M. Li, and A. J. Smola<br>
<a href="https://arxiv.org/abs/2210.03493" target="_blank">Automatic Chain of Thought Prompting in Large Language Models</a><br>
*"Let's think not just step by step, but also one by one."*
In *arXiv*, 2022<br>
<a href="https://twitter.com/astonzhangAZ/status/1579489453789581312" title="Tweet" target="_blank"><img src="./img/logo-twitter.svg" class="paper-icon"></a>
<a href="https://github.com/amazon-research/auto-cot" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* J. Chen, **A. Zhang**, X. Shi, M. Li, A. J. Smola, and D. Yang<br>
<a href="https://d2l-webdata.s3.us-west-2.amazonaws.com/papers/PEFT-design-spaces.pdf" target="_blank">Parameter-Efficient Fine-Tuning Design Spaces</a><br>
*"Adapters, prefix tuning, BitFit, LoRA, ... do design patterns exist?"*
In *arXiv*, 2022<br>

* M. S. Bari, **A. Zhang**, S. Zheng, X. Shi, Y. Zhu, S. Joty, M. Li<br>
<a href="https://arxiv.org/abs/2212.10929" target="_blank">
SPT: Semi-Parametric Prompt Tuning for Multitask Prompted Learning</a><br>
*"If your prompt tuning can't converge easily, make it semi-parametric."*
In *arXiv*, 2022<br>
<a href="https://twitter.com/astonzhangAZ/status/1605959819894349824" title="Tweet" target="_blank"><img src="./img/logo-twitter.svg" class="paper-icon"></a>

* H. Wang, **A. Zhang**, Y. Zhu, S. Zheng, M. Li, A. J. Smola, and Z. Wang<br>
<a href="https://arxiv.org/pdf/2207.01160.pdf" target="_blank">Partial and Asymmetric Contrastive Learning for Out-of-Distribution Detection in Long-Tailed Recognition</a><br>
In *Proceedings of International Conference on Machine Learning* (**ICML, Long Presentation**), 2022<br>
<a href="https://twitter.com/astonzhangAZ/status/1549800894840971265" title="Tweet" target="_blank"><img src="./img/logo-twitter.svg" class="paper-icon"></a>
<a href="https://github.com/amazon-research/long-tailed-ood-detection" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* H. Wang, **A. Zhang**, S. Zheng, X. Shi, M. Li, and Z. Wang<br>
<a href="https://arxiv.org/pdf/2207.01156.pdf" target="_blank">Removing Batch Normalization Boosts Adversarial Training</a><br>
In *Proceedings of International Conference on Machine Learning* (**ICML**), 2022<br>
<a href="https://twitter.com/astonzhangAZ/status/1549069378250874880" title="Tweet" target="_blank"><img src="./img/logo-twitter.svg" class="paper-icon"></a>
<a href="https://github.com/amazon-research/normalizer-free-robust-training" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* H. Wang, J. Hong, **A. Zhang**, J. Zhou, and Z. Wang<br>
<a href="https://arxiv.org/abs/2210.06428" target="_blank">Trap and Replace: Defending Backdoor Attacks by Trapping Them into an Easy-to-Replace Subnetwork</a><br>
In *Proceedings of the Conference on Neural Information Processing Systems* (**NeurIPS**), 2022

* Z. Hu, R. K.-W, Lee, C. C. Aggarwal, and **A. Zhang**<br>
<a href="https://arxiv.org/pdf/2010.12742.pdf" target="_blank">Text Style Transfer: A Review and Experimental Evaluation</a><br>
In *ACM SIGKDD Explorations*, 2022<br>

* C. He, S. Zheng, **A. Zhang**, G. Karypis, T. Chilimbi, M. Soltanolkotabi, S. Avestimehr<br>
<a href="https://arxiv.org/abs/2212.05191" target="_blank">
SMILE: Scaling Mixture-of-Experts with Efficient Bi-level Routing</a><br>
*"2.5x speedup over Switch Transformers"*
In *arXiv*, 2022<br>

* X. Hao, Y. Zhu, S. Appalaraju, **A. Zhang**, W. Zhang, B. Li, and M. Li<br>
<a href="https://arxiv.org/abs/2206.08358" target="_blank">MixGen: A New Multi-Modal Data Augmentation</a><br>
*"Interpolate images, concatenate text."*
In *arXiv*, 2022<br>
<a href="https://twitter.com/astonzhangAZ/status/1537860202652786688" title="Tweet" target="_blank"><img src="./img/logo-twitter.svg" class="paper-icon"></a>
<a href="https://github.com/amazon-research/mix-generation" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* E. Grassucci, **A. Zhang**, and D. Comminiello<br>
<a href="https://arxiv.org/abs/2110.04176" target="_blank">PHNNs: Lightweight Neural Networks via Parameterized Hypercomplex Convolutions</a><br>
In *IEEE Transactions on Neural Networks and Learning Systems* (**TNNLS**), 2022<br>
<a href="https://github.com/eleGAN23/HyperNets" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

## 2021

* **A. Zhang**, Y. Tay, S. Zhang, A. Chan, A. T. Luu, S. C. Hui, and J. Fu<br>
<a href="https://arxiv.org/abs/2102.08597" target="_blank">Beyond Fully-Connected Layers with Quaternions: Parameterization of Hypercomplex Multiplications with 1/n Parameters</a><br>
In *Proceedings of the International Conference on Learning Representations* (**ICLR, <font color="red">Outstanding Paper Award</font>**), 2021<br>
<a href="https://github.com/astonzhang/Parameterization-of-Hypercomplex-Multiplications" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* A. Chan, Y. S. Ong, B. Pung, **A. Zhang**, and J. Fu<br>
<a href="https://arxiv.org/pdf/2006.03535.pdf" target="_blank">CoCon: A Self-Supervised Approach for Controlled Text Generation</a><br>
In *Proceedings of the International Conference on Learning Representations* (**ICLR**), 2021<br>
<a href="https://github.com/alvinchangw/COCON_ICLR2021" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* **A. Zhang**, Y. Tay, Y. Shen, A. Chan, and S. Zhang<br>
<a href="https://proceedings.neurips.cc/paper/2021/file/3341f6f048384ec73a7ba2e77d2db48b-Paper.pdf" target="_blank">Self-Instantiated Recurrent Units with Dynamic Soft Recursion</a><br>
In *Proceedings of the Conference on Neural Information Processing Systems* (**NeurIPS**), 2021<br>
<a href="https://github.com/astonzhang/Self-IRU" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* Y. Long, B. Wang, Z. Yang, B. Kailkhura, **A. Zhang**, C. A. Gunter, and B. Li<br>
<a href="https://arxiv.org/pdf/1906.09338.pdf" target="_blank">G-PATE: Scalable Differentially Private Data Generator via Private Aggregation of Teacher Discriminators</a><br>
In *Proceedings of the Conference on Neural Information Processing Systems* (**NeurIPS**), 2021<br>
<a href="https://github.com/AI-secure/G-PATE" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* T. Chen, Y. Sui, X. Chen, **A. Zhang**, and Z. Wang<br>
<a href="https://arxiv.org/abs/2102.06790" target="_blank">A Unified Lottery Ticket Hypothesis for Graph Neural Networks</a><br>
In *Proceedings of International Conference on Machine Learning* (**ICML**), 2021<br>
<a href="https://github.com/VITA-Group/Unified-LTH-GNN" title="Code" target="_blank"><img src="./img/logo-github.svg" class="paper-icon"></a>

* H. Shao, Z. Xiao, S. Yao, D. Sun, **A. Zhang**, S. Liu, T. Wang, J. Li, and T. Abdelzaher<br>
**ControlVAE: Tuning, Analytical Properties, and Performance Analysis**<br>
In *Transactions on Pattern Analysis and Machine Intelligence* (**TPAMI**), 2021

* **A. Zhang**, A. Chan, Y. Tay, J. Fu, S. Wang, S. Zhang, H. Shao, S. Yao, and R. Lee<br>
**On Orthogonality Constraints for Transformers**<br>
In *Proceedings of the 59th Annual Meeting of the Association for Computational Linguistics* (**ACL, Oral**), 2021

* H. Shao, J. Wang, H. Lin, X. Zhang, **A. Zhang**, H. Ji, and T. Abdelzaher<br>
**Controllable and Diverse Text Generation in E-commerce**<br>
In *Proceedings of the Web Conference* (**WWW**), 2021

* S. Zhang, H. Liu, **A. Zhang**, Y. Hu, C. Zhang, Y. Li, T. Zhu, S. He, and W. Ou<br>
**Learning User Representations with Hypercuboids for Recommender Systems**<br>
In *Proceedings of the 14th International Conference on Web Search and Data Mining* (**WSDM**), 2021

## 2020

* H. Shao, S. Yao, D. Sun, **A. Zhang**, S. Liu, D. Liu, J. Wang, and T. Abdelzaher<br>
**ControlVAE: Controllable Variational Autoencoder**<br>
In *Proceedings of International Conference on Machine Learning* (**ICML**), 2020

* J. Guo, H. He, T. He, L. Lausen, M. Li, H. Lin, X. Shi, C. Wang, J. Xie, S. Zha, **A. Zhang**, H. Zhang, Z. Zhang, Z. Zhang, S. Zheng, and Y. Zhu<br>
**GluonCV and GluonNLP: Deep Learning in Computer Vision and Natural Language Processing**<br>
In *Journal of Machine Learning Research* (**JMLR**), Vol. 21, No. 23, 2020

* A. Chan, Y. Tay, Y. S. Ong, and **A. Zhang**<br>
**Poison Attacks against Text Datasets with Conditional Adversarially Regularized Autoencoder**<br>
In *Findings of Empirical Methods in Natural Language Processing* (**EMNLP-Findings**), 2020

* H. Shao, D. Sun, J. Wu, Z. Zhang, **A. Zhang**, S. Yao, S. Liu, T. Wang, C. Zhang, and T. Abdelzaher<br>
**GitHub Repository Recommendation for Academic Papers**<br>
In *Proceedings of the Web Conference* (**WWW**), 2020

## 2019

* Y. Tay, A. T. Luu, **A. Zhang**, S. Wang, and S. C. Hui<br>
**Compositional De-Attention Networks**<br>
In *Proceedings of the Conference on Neural Information Processing Systems* (**NeurIPS**), 2019

* S. Zhang, L. Yao, L. V. Tran, **A. Zhang**, and Y. Tay<br>
**Quaternion Collaborative Filtering for Recommendation**<br>
In *Proceedings of the 28th International Joint Conference on Artificial Intelligence* (**IJCAI**), 2019

* Y. Tay, **A. Zhang**, A. T. Luu, J. Rao, S. Zhang, S. Wang, J. Fu, and S. C. Hui<br>
**Lightweight and Efficient Neural Natural Language Processing with Quaternion Networks**<br>
In *Proceedings of the 57th Annual Meeting of the Association for Computational Linguistics* (**ACL**), 2019

* Y. Tay, S. Wang, A. T. Luu, J. Fu, M. C. Phan, X. Yuan, J. Rao, S. C. Hui, and **A. Zhang**<br>
**Simple and Effective Curriculum Pointer-Generator Networks for Reading Comprehension over Long Narratives**<br>
In *Proceedings of the 57th Annual Meeting of the Association for Computational Linguistics* (**ACL**), 2019

* A. El-Kishky, F. Xu, **A. Zhang**, and J. Han<br>
**Parsimonious Morpheme Segmentation with an Application to Enriching Word Embeddings**<br>
In *Proceedings of the IEEE International Conference on Big Data* (**IEEE BigData**), 2019

## Earlier

* S. Yao, Y. Zhao, **A. Zhang**, S. Hu, H. Shao, C. Zhang, L. Su, and T. Abdelzaher<br>
**Deep Learning for the Internet of Things**<br>
In *Computer*, 2018 (selected as May 2018 cover feature of the flagship magazine of the IEEE Computer Society)

* A. El-Kishky, F. Xu, **A. Zhang**, S. Macke, and J. Han<br>
**Entropy-Based Subword with an Application to Word Embeddings**<br>
In *Proceedings of the 2nd Workshop on Subword and Character Level Models in NLP at the 16th Annual Conference of the North American Chapter of the Association for Computational Linguistics: Human Language Technologies* (**NAACL SCLeM**), 2018

* S. Yao, Y. Zhao, H. Shao, C. Zhang, **A. Zhang**, S. Hu, D. Liu, S. Liu, L. Su, and T. F. Abdelzaher<br>
**SenseGAN: Enabling Deep Learning for Internet of Things with a Semi-Supervised Framework**<br>
In *Proceedings of the ACM on Interactive, Mobile, Wearable and Ubiquitous Technologies* (**ACM Ubicomp, <font color="red">Distinguished Paper Award</font>**), 2018

* **A. Zhang**, X. Lu, C. A. Gunter, S. Yao, F. Tao, R. Zhu, H. Gui, D. Fabbri, D. Liebovitz, and B. Malin<br>
**De Facto Diagnosis Specialties: Recognition and Discovery**<br>
In *Learning Health Systems* (**LHS**), 2018

* S. Yao, Y. Zhao, H. Shao, C. Zhang, **A. Zhang**, D. Liu, S. Liu, L. Su, and T. Abdelzaher<br>
**ApDeepSense: Deep Learning Uncertainty Estimation Without the Pain for IoT**<br>
In *Proceedings of the 38th IEEE International Conference on Distributed Computing Systems* (**ICDCS**), 2018

* **A. Zhang**, L. Garcia-Pueyo, J. B. Wendt, M. Najork, and A. Broder<br>
**Email Category Prediction**<br>
In *Proceedings of the 26th International World Wide Web Conference * (**WWW**), 2017

* S. Yao, S. Hu, Y. Zhao, **A. Zhang**, and T. Abdelzaher<br>
**DeepSense: A Unified Deep Learning Framework for Time-Series Mobile Sensing Data Processing**<br>
In *Proceedings of the 26th International World Wide Web Conference* (**WWW**), 2017

* S. Yao, Y. Zhao, **A. Zhang**, S. Lu, and T. Abdelzaher<br>
**DeepIoT: Compressing Deep Neural Network Structures for Sensing Systems without Loss of Performance**<br>
In *Proceedings of the 15th ACM Conference on Embedded Networked Sensor Systems* (**ACM Sensys, <font color="red">Best Paper Award Nomination</font>**), 2017

* S. Yao, Y. Zhao, H. Shao, **A. Zhang**, C. Zhang, S. Li, and T. Abdelzaher<br>
**RDeepSense: Reliable Deep Mobile Computing Models with Uncertainty Estimations**<br>
In *Proceedings of the ACM on Interactive, Mobile, Wearable and Ubiquitous Technologies* (**ACM Ubicomp**), 2017

* **A. Zhang**, A. Goyal, R. Baeza-Yates, Y. Chang, J. Han, C. A. Gunter, and H. Deng<br>
**Towards Mobile Query Auto-Completion: An Efficient Mobile Application-Aware Approach**<br>
In *Proceedings of the 25th International World Wide Web Conference* (**WWW**), 2016

* **A. Zhang** and Q. Gu<br>
**Accelerated Stochastic Block Coordinate Descent with Optimal Sampling** [<a href="papers/supp_asbcd.pdf" target="_blank">Proof of Lemmas</a>]<br>
In *Proceedings of the 22nd ACM SIGKDD Conference on Knowledge Discovery and Data Mining* (**KDD**), 2016

* S. Demetriou, W. Merrill, W. Yang, **A. Zhang**, and C. A. Gunter<br>
**Free for All! Assessing User Data Exposure to Advertising Libraries on Android**<br>
In *Proceedings of the Network and Distributed System Security Symposium* (**NDSS**), 2016

* R. Zhu, **A. Zhang**, J. Peng, and C. Zhai<br>
**Exploiting Temporal Divergence of Topic Distributions for Event Detection**<br>
In *Proceedings of the IEEE International Conference on Big Data* (**IEEE BigData**), 2016

* **A. Zhang**, A. Goyal, W. Kong, H. Deng, A. Dong, Y. Chang, C. A. Gunter, and J. Han<br>
**adaQAC: Adaptive Query Auto-Completion via Implicit Negative Feedback**<br>
In *Proceedings of the 38th International ACM SIGIR Conference on Research and Development in Information Retrieval* (**SIGIR**), 2015

* W. Kong, R. Li, J. Luo, **A. Zhang**, Y. Chang, and J. Allan<br>
**Predicting Search Intent Based on Pre-Search Context**<br>
In *Proceedings of the 38th International ACM SIGIR Conference on Research and Development in Information Retrieval* (**SIGIR**), 2015

* H. Fu, **A. Zhang**, and X. Xie<br>
**Effective Social Graph De-anonymization based on Graph Structure and Descriptive Information**<br>
In *ACM Transactions on Intelligent Systems and Technology* (**ACM TIST**), Vol. 6, No. 4, 2015

* X. Lu, **A. Zhang**, C. A. Gunter, D. Fabbri, D. Liebovitz, and B. Malin<br>
**Discovering De Facto Diagnosis Specialties**<br>
In *Proceedings of the 6th ACM Conference on Bioinformatics, Computational Biology, and Health Informatics* (**ACM BCB**), 2015

* **A. Zhang**, X. Xie, K. C.-C. Chang, C. A. Gunter, J. Han, and X. Wang<br>
**Privacy Risk in Anonymized Heterogeneous Information Networks**<br>
In *Proceedings of the 17th International Conference on Extending Database Technology* (**EDBT**), 2014

* H. Fu, **A. Zhang**, and X. Xie<br>
**De-anonymizing Social Graphs via Node Similarity**<br>
In *Proceedings of the 23rd International World Wide Web Conference* (**WWW**), 2014